---
title: "Santa Barbara Urchins"
author: "Roshni Katrak-Adefowora"
date: "10/31/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#Load packages
library(tidyverse)
library(here)
library(janitor)
library(kableExtra)
library(lubridate)
```

## Read in the urchins data
```{r}
urchins <- read_csv(here("data", "urchins.csv")) #read in data

urchins_clean <- urchins %>% 
  clean_names() #convert to lowercase snake case
```

## Compare sizes of purple and red urchins

Make exploratory graphs (faceted histogram and a boxplot) comparing the urchin sizes.
```{r, message = FALSE, warning = FALSE}
#Faceted histogram
ggplot(data=urchins_clean, aes(x=size))+ #size on x-axis
  geom_histogram()+ #histogram
  facet_wrap(~common_name)+ #two separate graphs for purple urchin and red urchin
  labs(x="Urchin diameter (cm)", y="Frequency") #update x-axis and y-axis

#Boxplot
ggplot(data=urchins_clean, aes(x=common_name, y=size))+
  geom_boxplot()+
  labs(x="Common Name", y="Urchin diameter (cm)")

```

Make summary statistics table for purple and red urchins
```{r, message = FALSE, warning = FALSE}
urchins_size <- urchins_clean %>% 
  group_by(common_name) %>% #group the purple urchins together and red urchins together
  summarize(mean_size_cm = mean(size, na.rm=TRUE), #add mean size column
          median_size_cm = median(size, na.rm=TRUE), #add median size column
          sd = sd(size, na.rm=TRUE), #add standard deviation column
          sample_size = n()) #add sample size column

#use kableExtra for better presentation
urchins_size %>%
kable(col.names = c("Common Name",
                      "Mean size (cm)", 
                      "Median size (cm)",
                      "Standard deviation",
                      "Sample size (n)")) %>% 
  kable_styling(full_width = FALSE)

#actual mean difference
6.776489-4.469797

#percent mean difference
(6.776489-4.469797)/((6.776489+4.469797)/2) * 100
```

Based on the graphs and summary table, red urchins on average are larger than purple urchins (2.31cm larger, 41.0 percent difference). The purple urchin has a smaller standard deviation, which indicates that the spread of values stays closer to the mean, whereas in the red urchin the data is more spread out (this can be seen in the histograms as well).

## Explore purple urchin counts by site and year

Subset data to only include purple urchins
```{r, message = FALSE, warning = FALSE}
purple_urchins <- urchins_clean %>% 
  filter(common_name == "Purple Urchin") %>% #filter to only include purple urchins
  mutate(date = mdy(date)) %>% #convert date to class "Date"
  mutate(year = year(date)) %>% #add column for year
  mutate(site_full = case_when(
    site == "NAPL" ~ "Naples Reef", #if site is NAPL, site_full is Naples Reef
    site == "AQUE" ~ "Arroyo Quemado Reef", #if site is AQUE, site_full is Arroyo Quemoado Reef
    site == "CARP" ~ "Carpinteria Reef", #if site is CARP, site_full is Carpinteria Reef
    site == "MOHK" ~ "Mohawk Reef", #if site is MOHK, site_full is Mohawk Reef
    site == "IVEE" ~ "Isla Vista Reef" #if site is IVEE, site_full is Isla Vista Reef
  ))
```

Make summary statistics table for purple urchins and plot the data
```{r, message = FALSE, warning = FALSE}
purple_sum <- purple_urchins %>% 
  count(site_full, year) #total number of purple urchins by site and year

ggplot(data=purple_sum, aes(x=year, y=n))+
  geom_col(aes(fill=site_full))+
  facet_wrap(~site_full)+
  labs(x="Year",
       y="Total number of purple urchins",
       fill="Site Name")+
  theme_bw()
```

