---
title: "Naples Reef (Jan 2010) Purple Urchins"
author: "Roshni Katrak-Adefowora"
date: "11/4/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
library(kableExtra)
library(here)
library(janitor)
```

## Read in the urchins data
```{r, message = FALSE, warning = FALSE}
urchins <- read_csv(here("data", "urchins.csv")) #read in data

urchins_clean <- urchins %>% 
  clean_names() #convert to lowercase snake case
```

## Filter to only include purple urchins at NAPL in January 2010
```{r, message = FALSE, warning = FALSE}
purple_urchins <- urchins_clean %>% 
  mutate(date = mdy(date)) %>% #convert date to class "Date"
  mutate(year = year(date)) %>%  #add column for year
  mutate(month = month(date)) %>% #add column for month
    filter(common_name == "Purple Urchin", site == "NAPL", month == "1", year == "2010") #purple urchins at NAPL in January 2010
```

## Exploratory histogram and QQ plots
```{r, message = FALSE, warning = FALSE}
#Faceted histogram
ggplot(data=purple_urchins, aes(x=size))+
  geom_histogram(bins=4)+ #4 bins
  facet_wrap(~treatment)+ #separate graphs for annual and control treatments
  labs(x="Urchin diameter (cm)", y="Frequency") #update x-axis and y-axis

#QQ-plot
ggplot(data=purple_urchins, aes(sample=size))+
  geom_qq()+
  facet_wrap(~treatment) #separate graphs for annual and control treatments
```

The data under both annual and control treatments appear to be somewhat normally distributed in the histogram (bell-shaped), and the qq-plot shows a mostly linear relationship which also indicates normality.

## Table of summary statistics
```{r, message = FALSE, warning = FALSE}
purple_size <- purple_urchins %>% 
  group_by(treatment) %>% #group annuals together and controls together
  summarize(mean_size_cm = mean(size, na.rm = TRUE), #add mean column
            sd = sd(size, na.rm = TRUE), #add standard deviation column
            sample_size = n()) #add sample size column

#use kableExtra for better presentation
purple_size %>% 
  kable(col.names = c("Treatment",
                      "Mean size (cm)",
                      "Standard deviation",
                      "Sample size (n)")) %>% 
  kable_styling(full_width = FALSE)
```

## Two-sample t-test
```{r, message = FALSE, warning = FALSE}
#Create vectors containing observations for each treatment
annual <- purple_urchins %>% 
  filter(treatment == "ANNUAL") %>% 
  pull(size) #turn dataframe into vector

control <- purple_urchins %>% 
  filter(treatment == "CONTROL") %>% 
  pull(size) #turn dataframe into vector

#Two-sample t-test
t.test(annual, control)
```

